\newpage
\chapter{Code Structure}

The code repository can be found at: \url{https://github.com/AceMouse/xTB-math}.

\dirtree{%
.1 xTB-math/.
.2 bin2xyz/.
.3 bin2xyz.cpp\DTcomment{converter from float64 coords into xyz format}.
.3 C200\_10000\_fullerenes.float64\DTcomment{3d coords for 10k fullerenes}.
.2 flake.lock.
.2 flake.nix\DTcomment{conventional structure for Nix inputs and outputs}.
.2 nix/\DTcomment{Nix package definitions for xtb and its dependencies}.
.3 cpx.nix\DTcomment{CPCM-X - a solvation model}.
.3 numsa.nix\DTcomment{solvent accessible surface area calculation}.
.3 patches/\DTcomment{patches for extracting data for validation}.
.4 dftd4/\DTcomment{dispersion correction}.
.5 log\_args\_and\_outputs.patch.
.5 use\_gfn2.patch.
.4 xtb/.
.5 log\_args\_and\_outputs.patch.
.5 log\_electro.patch.
.5 log\_utils.patch.
.3 xtb.nix\DTcomment{extended tight-binding program}.
.2 README.md.
.2 report/.
.2 xtb-python/\DTcomment{Python port of xTB paper and Fortran impl}.
.3 basisset.py.
.3 blas.py.
.3 cmp\_impls.py\DTcomment{validation tests against Fortran impl}.
.3 data/.
.4 C200.xyz.
.4 caffeine.xyz.
.3 dftd4.py\DTcomment{computation for dispersion correction}.
.3 dftd4\_reference.py\DTcomment{constants for dftd4}.
.3 energy.py\DTcomment{various energy computations}.
.3 fock.py\DTcomment{Fock matrix computation}.
.3 gfn2.py\DTcomment{GFN2-xTB specific constants}.
.3 lapack.py\DTcomment{Facade for LAPACK functions}.
.3 scc.py\DTcomment{computation for self-consistent charges}.
.3 slater.py\DTcomment{computation for slater determinants}.
.3 util.py.
.3 xyz\_reader.py.
}

\newpage

The code repository can be found at: \url{https://github.com/Maroka-chan/xtb-gpu}.
\dirtree{%
.1 xtb-gpu/.
.2 flake.lock.
.2 flake.nix.
.2 nix/.
.3 nvhpc.nix\DTcomment{patching nvhpc to make nvfortran work on Nix}.
.3 xtb.nix\DTcomment{xtb version 6.4.0 compiled with nvfortran}.
.2 README.md.
.2 sycl/.
.3 build\_SDQH0.cpp\DTcomment{incomplete SYCL impl for build\_SDQH0}.
.3 data\_caffeine/\DTcomment{test data for electro.cpp computed from caffeine}.
.4 atomicGam.txt.
.4 dqsh.txt.
.4 dq.txt.
.4 H0.txt.
.4 jmat.txt.
.4 P.txt.
.4 shellGam.txt.
.4 shift.txt.
.3 data\_c200/\DTcomment{test data for electro.cpp computed from a C200 fullerene}.
.4 atomicGam.txt.
.4 ....
.3 electro.cpp\DTcomment{SYCL impl for computing electrostatic energy}.
}



