\documentclass{article}
\usepackage{amsmath}
\usepackage{braket}

\newcommand{\q}[1]{\textbf{QUESTION: #1}}
\newcommand{\A}[1]{\textbf{ANSWER: #1}}
\makeatletter
\newcommand\E{\@ifnextchar[{\@with}{\@without}}
\def\@with[#1]#2{E_{#2}^{(#1)}}
\def\@without#1{E_{#1}}
\makeatother

\makeatletter
\newcommand\footnoteref[1]{\protected@xdef\@thefnmark{\ref{#1}}\@footnotemark}
\makeatother

\begin{document}

\section{Gaussian functions per reference system}
\noindent
The Gaussian weighting for each reference system is given by:
\begin{equation}
  W_A^r(CN_{cov}^A, CN_{cov}^{A,r}) = \sum_{j=1}^{N_{gauss}} \frac{1}{\mathcal{N}} \exp\left[-6j \cdot (CN_{cov}^A - CN_{cov}^{A,r})^2\right]
\end{equation}

\noindent
with
\begin{equation}
  \sum_{r}^{N_{A,ref}} W_A^r(CN_{cov}^A, CN_{cov}^{A,r}) = 1
\end{equation}

\vspace{10pt}
\noindent
\(\mathcal{N}\) is a normalization constant.

\noindent
The number of Gaussian functions per reference system \(N_{gauss}\) is mostly one, but equal to three for \(CN_{cov}^{A,r} = 0\) and reference systems with similar coordination number.

\q{How do we obtain the number of gaussian functions per reference system (\(N_{gauss})\)?}

\q{How or where do we get the normalization constant?}



\section{SAD - Superposition of Atomic Densities}

The superposition of atomic densities(SAD) is an approach to obtain a good approximation of a collection of atoms, to be used as an initial guess for solving the self-consistent field(SCF) equation.

As originally implemented in DISCO, the molecular electron density can be obtained by adding the densities of all the constituting atoms.


%https://pyscf.org/user/scf.html#initial-guess
%https://sci-hub.box/10.1002/jcc.20393

This is how we get the density matrix for an isolated atom?

\begin{equation}
  \rho_0 = \sum_A \rho_0^A
\end{equation}
\begin{equation}
  D_{ij} = \sum_a^{occ} c_{ia} c_{ja}
\end{equation}

To get the coefficients we need to solve SCF for each atom.
The SAD method is then the sum of all of these.

\vspace{10pt}
\noindent
\q{How can we get the initial guess for the Fock matrix and Coefficient matrices needed to run SCF on the individual atoms?}

Direct SCF Approach
\begin{equation}
\begin{split}
  \Delta F_{ab} = &(c_{ia}c_{jb} + c_{ja}c_{ib})\\
  &\Delta F_{ij} + (c_{ia}c_{kb} + c_{ka}c_{ib})\\
  &\Delta F_{ik} + (c_{ia}c_{lb} + c_{la}c_{ib})\\
  &\Delta F_{il} + (c_{ja}c_{kb} + c_{ka}c_{jb})\\
  &\Delta F_{jk} + (c_{ja}c_{lb} + c_{la}c_{jb})\\
  &\Delta F_{jl} + (c_{ka}c_{lb} + c_{la}c_{kb}) \Delta F_{kl}\\
  &= l_{ijkl}(4E_{ij}^{ab}D_{kl} + 4D_{ij}E_{kl}^{ab} - E_{ik}^{ab}D_{jl} - D_{ik}E_{jl}^{ab} - E_{il}^{ab}D_{jk} - D_{il}E_{jk}^{ab})
\end{split}
\end{equation}
where
\begin{equation}
  E_{ij}^{ab} = c_{ia}c_{jb} + c_{ja}c_{ib}
\end{equation}



\section{Questions for Albert}

\q{Does the xTB program compute the self-consistent field(SCF) or is that part of an external project?}

\vspace{10pt}
\noindent
\q{How do you get your initial guess for the Fock matrix and Coefficient matrices needed to run SCF on the individual atoms?}

\vspace{10pt}
\noindent
\q{Can you shed some light on where and how the xTB program computes the initial guess?}

\vspace{10pt}
\noindent
\q{How and where is the overlap matrix computed? Do you do the integrations yourself?}

\vspace{10pt}
\noindent
\q{For the dispersion term, how do we get \(\overline{\alpha}_A\) and \(\overline{\alpha}_{A,r}\)?}

\vspace{10pt}
\noindent
\q{For the dispersion term, how do we get the integration weights \(\omega\)?}

\vspace{10pt}
\noindent
\q{How do we get the atomic charges? Is it the mulliken charges?}


\A{yas, check equation 24 in main}

\vspace{10pt}
\noindent
\q{How do we get the covalent coordination number \(CN_\text{cov}^A\) and \(CN_\text{cov}^{A,r}\) for an atom A?}

\vspace{10pt}
\noindent
\q{What is \(N_{A,\text{ref}}\)? The number of reference... orbitals? for an atom A?}


\newpage
\noindent
\(\theta_{ABC}\) is the angle between the two edges going from B to the other two atoms.

\noindent
\q{Is this definition correct? I assume we have the coordinates for each atom in (x,y,z)?}

\A{%partial answer to the assumption only
    We do have the positions of each atom in terms of (x,y,z)}

\vspace{10pt}
\noindent
\q{In the Anisotropic Exchange-correlation energy term $E_{AES} = \sum_A(f_{XC}^{\mu_A}|\pmb\mu_A|^2 + f_{XC}^{\Theta_A}||\pmb\Theta_A||^2)$ Which norms are computed? $|\cdot |^2$ is presumably the Euclidean norm, how about the matrix norm?}

\vspace{10pt}
\noindent
\q{If $\alpha$ is a cartesian component say the label $x$, is $\alpha_i$ then the value associated with that label? (see eq. 27b,c in the GFN2 paper)}

\vspace{10pt}
\noindent
\q{For calculating $V_Q^{\alpha\beta}$ in eq. 44 in the GFN2 paper, which B are we using in the $R_{AB}^2$ term? (should it be a C?)}\\
\q{$\alpha_{AC}^2$ is the cartesian component $\alpha$ in the $R_{AC}$ vector, correct? (see eq. 44 in the GFN2 paper)}
\begin{align*}
    \begin{split}
        V_Q^{\alpha\beta}(\pmb{R}_C)
            &= -\sum_A q_Af_5(R_{AC})\left[          
                \frac{3}{2}\alpha_{AC}\beta_{AC}- \frac{1}{2}R_{AB}^2 % TODO: is AB correct here!?
                \right]\\
        &-f_{XC}^{\Theta_C}\left[3\pmb\Theta_C^{\alpha\beta}-\delta_{\alpha\beta}\sum_\alpha\pmb\Theta_C^{\alpha\alpha}\right]
    \end{split}
\end{align*}
\q{Is the dispersion constant $s_{42} = 0.5$? (it seems this way in the code, but is not mentioned anywhere.)}

\end{document}
